---
title: "BIOL432 Assignment 5"
author: "SiruiZHAO"
date: "2023-02-08"
output: html_document
---

Part I: Data Science

[my GitHub repository](https://github.com/siruiZHAO/BIOL432_Assignment_5.git)

```{r}
#load library
library(dplyr)
```

1. Import your measurements.csv file from the Week 1 Assignment into an object called MData
```{r}
#Import data
MData <- read.csv("measurements.csv")

head(MData)
```
```{r}
#check data structure
str(MData)
```
    
2. Use a dplyr command AND regular expression(s) to add a new column called Sp that shortens the genus name in the Organism column to only the first letter and period. For example: Homo sapiens in the original Organism column becomes H. sapiens in the Sp column.
    
```{r}
#use mutate() create new column and gsub() in regular expressions to shortens the letters after first capital letters
MData <- MData %>%
  mutate(Sp = gsub("([A-Z]).* (.*)", "\\1. \\2", organism))

#check data
head(MData)
```
    
3. Use a dplyr command to create a new dataset BUT with Sp column instead of the Organism column, moved to the beginning in place of the original Organism column.
```{r}
#create new dataset only with Sp
#move Sp column to the beginning
NewData <- MData %>%
  select(Sp, Limb_width, Limb_length, limb_volume, Observer)

#check data
head(NewData)
```

4. Save the dataset as a new file called MeasShort.csv
```{r}
#save as csv file
write.csv(NewData, "MeasShort.csv")
```

Thanks!